Title Goes Here

### 2023-01-29 inst/paper/_materials.qmd
add: co-author feedback

<!-- Demographics and description of MB11 sample, selection criteria. -->
The sample consists of `r nrow(demography)` individuals from Middenbeemster,
a 19th century rural Dutch site. 
The village of Middenbeemster and the surrounding Beemsterpolder was established
in the beginning of the 17th century, when the Beemster lake was drained to
create more farm land, mainly for the cultivation of cole seeds (de Vries 1978).
In 1615, a decision was made to dedicate a church to Hendrick de Keyser, and
construction started in 1618 (Hakvoort 2013). The excavated cemetery is associated
with the Keyserkerk church, where the inhabitants of the Middenbeemster village
and the surrounding Beemsterpolder were buried between AD 1615 and 1866
[@lemmersMiddenbeemster2013]. Archival documents are available for those buried
between AD 1829 and 1866, when the majority of individuals were interred
[@palmerActivityReconstruction2016]. The main occupation of the inhabitants
was dairy farming, consisting largely of manual labour prior to the industrial
revolution [@aten400JaarBeemster2012; @palmerActivityReconstruction2016].

To reduce the number of factors to account for
in the analysis, we preferentially selected males from the middle adult age category
(35-49). The sample consists of `r nrow(filter(demography, sex == "m"))` males,
`r nrow(filter(demography, sex == "pm"))` probable males,
`r nrow(filter(demography, sex == "pf"))` probable females, and
`r nrow(filter(demography, sex == "f"))` female [@fig-sample-demography].
We selected males due to a higher
occurrence of dental calculus deposits than females (unpublished observation).

<!-- use plot to show original and replicated sample demographics? -->
```{r}
#| label: fig-sample-demography
#| fig-cap: "Overview of sample demography. Left plot is the first batch and right plot is replication batch. eya = early young adult (18-24); lya = late young adult (25-34); ma = middle adult (35-49); old = old adult (50+)."

>>>>

<!-- Demographics and description of MB11 sample, selection criteria. -->
The sample consists of `r nrow(demography)` individuals from Middenbeemster,
a 19th century rural Dutch site.
Of the `r nrow(demography)` individuals that were analysed,
`r nrow(filter(metadata, replicated == TRUE))` individuals were subsequently
replicated using leftover calculus from the original DNA study.
<!--in the same lab? different lab? -->
The village of Middenbeemster and the surrounding Beemsterpolder was established
in the beginning of the 17th century, when the Beemster lake was drained to
create more farm land, mainly for the cultivation of cole seeds (de Vries 1978).
In 1615, a decision was made to build a church, and
construction started in 1618 (Hakvoort 2013). The excavated cemetery is associated
with the Keyserkerk church, where the inhabitants of the Middenbeemster village
and the surrounding Beemsterpolder were buried between AD 1615 and 1866
[@lemmersMiddenbeemster2013]. Archival documents are available for those buried
between AD 1829 and 1866, when the majority of individuals were interred
[@palmerActivityReconstruction2016]. The main occupation of the inhabitants
was dairy farming, consisting largely of manual labour prior to the industrial
revolution [@aten400JaarBeemster2012; @palmerActivityReconstruction2016].

To reduce the number of factors to account for
in the analysis, we preferentially selected males from the middle adult age category
(35-49 years). The sample consists of `r nrow(filter(demography, sex == "m"))` males,
`r nrow(filter(demography, sex == "pm"))` probable males,
`r nrow(filter(demography, sex == "pf"))` probable females, and
`r nrow(filter(demography, sex == "f"))` female [@fig-sample-demography].
We selected males due to a higher
occurrence of dental calculus deposits than females (unpublished observation).

<!-- use plot to show original and replicated sample demographics? -->
```{r}
#| label: fig-sample-demography
#| fig-cap: "Overview of sample demography. Left plot is the first batch and right plot is the replication batch with 29 of the individuals from the first batch. eya = early young adult (18-24 years); lya = late young adult (25-34 years); ma = middle adult (35-49 years); old = old adult (50+ years). Male? = probable male; Female? = probable female."

### 2023-01-29 inst/paper/_methods.qmd
add: co-author feedback

New dental calculus samples were taken under sterile conditions in a positive
pressure laminar flow hood in a dedicated dental calculus lab at Leiden University.
The surface of the tooth was lightly brushed with a sterile, disposable toothbrush
to get rid of surface contaminants. A sterile dental curette was then used to
scrape calculus from the tooth onto weighing paper, which was transferred to 1.5 ml
Eppendorf tubes.
Where possible, previously sampled material was used to prevent repeated sampling
of individuals. The sampling methods can be found in the published material
[@ziesemer16SChallenges2015; @ziesemerGenomeCalculus2018]. 
The samples were sent to the Department of Forensic Medicine at Aarhus University
for ultra-high-performance liquid chromatography-tandem mass spectrometry method
using pneumatically assisted electrospray ionisation (UHPLC-ESI-MS/MS).

The UHPLC-ESI-MS/MS method was validated in a separate study on cadavers received
for forensic autopsy and toxicological analysis. Results from dental calculus were
validated against compounds detected in whole blood samples from the same
individuals.  
In the original method, samples were washed three times to remove residual
substances from the surface of the calculus that originated from oral fluids,
and only extract substances from the calculus. In our samples the washes served
to remove potential contaminants from the burial environment and post-excavation
handling.  
The list of targeted compounds included both naturally occurring compounds known
to have been used in the past, as well as synthetic modern drugs that did not exist
at the time (e.g. Fentanyl, MDMA, Amphetamine). These were part of the toxicology
screening for the original method [@sorensenDrugsCalculus2021],
developed on cadavers. In our study they serve as an authentication step, as their
presence in archaeological samples could only be the result of contamination.

Briefly, dental calculus was treated with citric acid and the dissolution extracts
were cleaned using weak and strong polymeric cation-exchange sorbents. Samples
were washed with 0.5 mL MeOH for 10 seconds. Samples were weighed before and after
each wash. Samples were air-dried for 24 hours at room temperature after each wash.
Extracts from each wash were analysed by injecting 5 mL volume into the column
of an Exion UHPLC system that consisted of two Exion AD pumps, an Exion AD
multiplate autosampler set at 10 $\pm$ 2 &deg;C and an Exion AC column oven set
at 40 $\pm$ 2 &deg;C (Sciex, Ontario, Canada).
Separation was performed using a Raptor Biphenyl UHPLC column (2.7 mm, 2.1 mm I.D.
$\times$100 mm) (Restek, Bellefonte, PA). The mass spectrometer was a Sciex QTRAP
6500+ with a TurboIonSpray probe for electrospray ionisation.  
The remaining calculus was dissolved using lysing tube beads in 800 $\mu$L of 0.5
$\small{M}$ citric acid (CA) and 50 $\mu$L stable isotope-labelled analogue used
as internal standards (SIL-IS) solution for 1 h at ambient
temperature with gentle shaking. The suspension was then mixed with 800 $\mu$L
MeOH and centrifuged at 10,000 $\times$ g for 5 mins, and analysed by the same
method as the wash extracts.
Data analysis was performed using Analyst 1.7 and MultiQuant 3.0.3 (Sciex).
Raw quantities of compounds are presented in ng and concentrations as ng / mg.  
For more details about the method and validation, see the original study by
Sørensen and colleagues [-@sorensenDrugsCalculus2021].

Of the `r nrow(demography)` individuals that were analysed,
`r nrow(filter(metadata, replicated == TRUE))` individuals were subsequently
replicated using leftover calculus from the original DNA study.
<!--in the same lab? different lab? -->

@@

<!-- methods for age, sex, and stature estimation -->
Demographic and pathological analyses were conducted in the Laboratory for Human
Osteoarchaeology at Leiden University. Sex was estimated using cranial
and pelvic morphological traits [@Standards1994]. Age-at-death was estimated using
dental wear,
auricular and pubic surface appearance, cranial suture closure, and epiphyseal fusion 
[@Standards1994; @SucheyBrooks1990; @buckberryAuricular2002; @lovejoyAuricular1985; @meindlSutureClosure1985], and divided into the following categories:
early young adult (EYA, 18-24), late young adult (LYA, 25-34), middle adult (MA, 
35-49), old adult (Old, 50+).

@@

Pathological conditions and lesions that occur frequently in the population were
included in the analysis. Data were
dichotomised to presence/absence to allow statistical analysis.
Osteoarthritis (OA) was considered present in cases where eburnation was visible
on one or more joint surface.
Vertebral osteophytosis is identified by marginal lipping and/or osteophyte
formation on the margin of the superior and inferior surfaces of the vertebral
body.
Cribra orbitalia (CO) was diagnosed based on presence of pitting on the superior
surface of the orbit. No distinction was made between active or <!-- ? --> lesions.
Degenerative disc disease (DDD), or spondylosis, is identified as a large diffuse
depression of the
superior and/or inferior surfaces of the vertebral body [@rogersPalaeopathologyJoint2000].
Schmorl's nodes (SN) are identified as any cortical depressions on the surface of
the vertebral body.
Data on chronic maxillary sinusitis from @casnaUrbanizationRespiratory2021 were
included in this study to assess the relationship between upper respiratory
diseases with environmental factors (i.e. tobacco smoke, caffeine consumption).
Lesions associated with chronic maxillary sinusitis (CMS) as defined by
@boocockMaxillarySinusitis1995 were
recorded for each individual and classified as "pitting", "spicule-type bone
formation", "remodeled spicules", or "white pitted bone". CMS was scored as absent
when the sinus presented smooth surfaces with little or no associated pitting.

>>>>

New dental calculus samples were taken under sterile conditions in a positive
pressure laminar flow hood in a dedicated dental calculus lab at Leiden University.
The surface of the tooth was lightly brushed with a sterile, disposable toothbrush
to get rid of surface contaminants. A sterile dental curette was then used to
scrape calculus from the tooth onto weighing paper, which was transferred to 1.5 ml
Eppendorf tubes.
Where possible, previously sampled material was used to prevent repeated sampling
of individuals. The sampling methods can be found in the published material
[@ziesemer16SChallenges2015; @ziesemerGenomeCalculus2018]. 
The samples were sent to the Department of Forensic Medicine at Aarhus University
for ultra-high-performance liquid chromatography-tandem mass spectrometry (UHPLC-MS/MS)
analysis.

The list of targeted compounds included both naturally occurring compounds known
to have been used in the past, as well as synthetic modern drugs that did not exist
at the time (e.g. Fentanyl, MDMA, Amphetamine). These were part of the toxicology
screening for the original method [@sorensenDrugsCalculus2021],
developed on cadavers. In our study they serve as an authentication step, as their
presence in archaeological samples could only be the result of contamination.

Briefly, samples of dental calculus were washed three times each with one mL of
methanol (MeOH), to remove surface contaminants. The wash solutions were collected
separately. The solvent was
evaporated and the residues were dissolved in 50 µL 30% MeOH.
The washed calculus was homogenized in presence of 0.5 M citric acid using a lysing
tube with stainless steel beads. Following one hour of incubation the dissolution
extract was cleaned by weak and strong cation-exchange. After evaporation of the
elution solvent the residue was dissolved in 50 µL 30% MeOH.
The final extracts obtained from washing and dissolution of the dental calculus were
analysed by UHPLC-MS/MS using a reversed-phase biphenyl column for chromatography.
To obtain quantitative results, isotope dilution was applied.
For more details about the method and validation, see the original study by
Sørensen and colleagues [-@sorensenDrugsCalculus2021].

@@

<!-- methods for age, sex, and stature estimation -->
Demographic and pathological analyses were conducted in the Laboratory for Human
Osteoarchaeology at Leiden University. Sex was estimated using cranial
and pelvic morphological traits [@Standards1994]. Age-at-death was estimated using
dental wear,
auricular and pubic surface appearance, cranial suture closure, and epiphyseal fusion 
[@Standards1994; @SucheyBrooks1990; @buckberryAuricular2002; @lovejoyAuricular1985; @meindlSutureClosure1985], and divided into the following categories:
early young adult (EYA, 18-24 years), late young adult (LYA, 25-34 years), middle adult (MA, 
35-49 years), old adult (Old, 50+ years).

@@

Pathological conditions and lesions that occur frequently in the population were
included in the analysis. Data were
dichotomised to presence/absence to allow statistical analysis.
Osteoarthritis (OA) was considered present in cases where eburnation was visible
on one or more joint surfaces.
Vertebral osteophytosis is identified by marginal lipping and/or osteophyte
formation on the margin of the superior and inferior surfaces of the vertebral
body.
Cribra orbitalia (CO) was diagnosed based on presence of pitting on the superior
surface of the orbit. No distinction was made between active or healing lesions.
Degenerative disc disease (DDD), or spondylosis, is identified as a large diffuse
depression of the
superior and/or inferior surfaces of the vertebral body [@rogersPalaeopathologyJoint2000].
Schmorl's nodes (SN) are identified as any cortical depressions on the surface of
the vertebral body.
Data on chronic maxillary sinusitis (CMS) from @casnaUrbanizationRespiratory2021 were
included in this study to assess the relationship between upper respiratory
diseases with environmental factors (i.e. tobacco smoke, caffeine consumption).
Lesions associated with CMS as defined by
@boocockMaxillarySinusitis1995 were
recorded for each individual and classified as "pitting", "spicule-type bone
formation", "remodeled spicules", or "white pitted bone". CMS was scored as absent
when the sinus presented smooth surfaces with little or no associated pitting.

### 2023-01-29 inst/paper/_results.qmd
add: co-author feedback

<!-- detected compounds -->
Multiple compounds were detected in the dental calculus samples. Compounds detected
at a lower concentration than the LLOQ were considered not present. Not all the
compounds detected in the first batch could be replicated in the second batch
(@tbl-compound-detect). For a full list of targeted compounds, see Supplementary
Material.

@@

<!-- May need to limit the list to compounds that were detected in either batch and keep the full list in supplementary materials -->
```{r}
#| label: tbl-compound-detect
#| tbl-cap: "Target compound including whether it was detected (TRUE) or not (FALSE) in each batch, as well as the lower limit of quantitation (LLOQ) in ng. CBD = cannabidiol; CBN = cannabinol; THC = tetrahydrocannabinol; THCA-A = tetrahydrocannabinolic acid; THCVA = tetrahydrocannabivarin acid."
uhplc_calculus_detect %>%
  left_join(compound_name_abbrev, by = c("compound" = "abbrev")) %>% 
  full_join(lloq, by = c("name" = "compound")) %>%
  mutate(across(batch1:batch2, ~ replace_na(.x, FALSE))) %>%
  filter(!(batch1 == FALSE & batch2 == FALSE)) %>% 
  select(name, batch1, batch2, lloq) %>% 
  arrange(name) %>% 
  knitr::kable(col.names = c("Compound", "Batch 1", "Batch 2", "LLOQ"))
```

@@

```{r}
#| label: fig-auth-plot-batch2
#| fig-cap: "(A) Number of samples in which the each compound was detected in the first and second batch. (B) Quantity (ng) of each compound extracted from each sample across the three washes and final calculus extraction (calc). Colours represent different calculus samples. Only results from batch 2 are shown."
#| fig-width: 8
#| fig-height: 7

@@

<!-- detection vs. preservation -->
To see if preservation of the skeletal remains had any effect on the detection of
compounds, concentrations of compounds were compared to the various levels of
preservation. No apparent patterns were present between preservation and detection
of compounds in the samples (@fig-detection-preservation).

```{r}
#| label: fig-detection-preservation
#| fig-cap: "(A) Distribution of extracted quantities of each compound, separated by state of preservation of the skeleton. (B) Extracted quantity (ng) of compound plotted against weights of the calculus samples from batch 2."
#| fig-height: 6
#| fig-width: 7

preservation_plot <- uhplc_calculus_long %>%
  filter(
    !is.na(preservation),
    !compound %in% c("cbd", "cbn", "cocaine", "thc", "thca-a", "thcva"), # remove compounds not detected
    id != "MB329" # unusually high nicotine and cotinine quantity in batch 1
    ) %>% 
  ggplot(aes(x = preservation, y = quant)) +
    geom_violin(aes(fill = preservation), alpha = 0.6) +
    #geom_jitter(width = 0.2) +
    geom_boxplot(width = 0.2) +
    facet_wrap(~ compound, scales = "free_y", labeller = labeller(compound = compound_names), dir = "v") +
    theme_bw() +
    theme(
      legend.position = "none",
      panel.grid.major.x = element_blank(),
      axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1)
      ) +
    labs(x = "Preservation", y = "Quantity (ng)") +
    scale_fill_viridis_d()

weight_quant_plot <- uhplc_data_comb %>%
  select(sample, contains("batch2")) %>% 
  select(sample, batch2_weight, contains("calc")) %>% 
  pivot_longer(
    -c(sample, batch2_weight),
    names_to = c("compound", "batch"),
    names_pattern = "(.*)_(.*)",
    values_to = c("conc")
  ) %>% 
  mutate(compound = str_remove(compound, "_calc")) %>%
  #remove_missing() %>% 
  filter(conc > 0) %>% 
  ggplot(aes(x = batch2_weight, y = conc)) +
    geom_point(col = viridisLite::viridis(1)) +
    facet_wrap(~ compound, scales = "free_y", labeller = labeller(compound = compound_names), dir = "v") +
    theme_bw() +
    theme(axis.title.y = element_blank()) +
    labs(y = "Quantity (ng)", x = "Weight (mg)")

@@

For further statistical analyses, only the UHPLC-ESI-MS/MS results from batch 2
were used, as batch 1 had multiple compounds that were not detected in batch 2
and may have been contamination.

```{r}
#| label: tbl-pearson
#| tbl-cap: "Pearson correlation (*r*) on dichotomous skeletal lesions and compound concentrations from the second batch. Correlations between pairs of dichotomous variables are removed due to incompatibility with a Pearson correlation. OA = osteoarthritis; VOP = vertebral osteophytosis; SN = Schmorl's nodes; DDD = degenerative disc disease; CO = cribra orbitalia; CMS = chronic maxillary sinusitis; SA = salicylic acid; PN = pipe notches."

@@

The strongest point-biserial (Pearson) correlation correlations were a near-perfect
positive correlation between `r cor_statement(corr_tib_long, strength = "strong")`,
and moderate correlations between
`r cor_statement(corr_tib_long, strength = "moderate")` (@tbl-pearson).

Polychoric correlation was conducted on the dichotomised compounds and pathological
conditions, as well as the discretised dental diseases.
Salicylic acid was removed due to its ubiquitous presence in the sample, and is
likely to cause spurious correlations.
Strong correlations were found between `r cor_statement(polycorr_tib, strength = "strong")`.
Moderate correlations were found between `r cor_statement(polycorr_tib, strength = "moderate")`.
Remaining correlations were weak or absent (@fig-polycorr).
Correlations with age will be depressed because age was largely controlled for
in the sample selection.

>>>>

<!-- detected compounds -->
Multiple compounds were detected in the dental calculus samples. Compounds detected
at a lower concentration than the lower limit of quantitation (LLOQ) were considered
not present. Not all the compounds detected in the first batch could be replicated
in the second batch (@tbl-compound-detect).
For a full list of targeted compounds, see Supplementary Material.

@@

<!-- May need to limit the list to compounds that were detected in either batch and keep the full list in supplementary materials -->
```{r}
#| label: tbl-compound-detect
#| tbl-cap: "Target compound including whether it was detected (TRUE) or not (FALSE) in each batch, as well as the lower limit of quantitation (LLOQ) in ng. CBD = cannabidiol; CBN = cannabinol; THC = tetrahydrocannabinol; THCA-A = tetrahydrocannabinolic acid A; THCVA = tetrahydrocannabivarin acid."
uhplc_calculus_detect %>%
  left_join(compound_name_abbrev, by = c("compound" = "abbrev")) %>% 
  full_join(lloq, by = c("name" = "compound")) %>%
  mutate(across(batch1:batch2, ~ replace_na(.x, FALSE))) %>%
  filter(!(batch1 == FALSE & batch2 == FALSE)) %>% 
  select(name, batch1, batch2, lloq) %>% 
  arrange(name) %>% 
  knitr::kable(col.names = c("Compound", "Batch 1", "Batch 2", "LLOQ"))
```

@@

```{r}
#| label: fig-auth-plot-batch2
#| fig-cap: "(A) Number of samples in which each compound was detected in the first and second batch. (B) Quantity (ng) of each compound extracted from each sample in batch 2. The plot displays the extracted quantity across the three washes and final calculus extraction (calc). Each coloured line represents a different calculus sample."
#| fig-width: 8
#| fig-height: 6

@@

<!-- detection vs. preservation -->
To see if preservation of the skeletal remains had any effect on the detection of
compounds, we compare extracted quantities of compounds to the various levels of
skeletal preservation. Our results from batch 2 suggest that detection of a compound
may be linked to the preservation of the skeleton, with better preservation leading
to increased extraction quantity ([@fig-detection-preservation]A). We also
find a weak positive correlation between the weight of the calculus sample and the
quantity of compound extracted from the calculus ([@fig-detection-preservation]B).

```{r}
#| label: fig-detection-preservation
#| fig-cap: "(A) Violin plot with overlaid box plots depicting the distribution of extracted quantities of each compound from batch 2 separated by state of preservation of the skeleton. (B) Extracted quantity (ng) of compound plotted against weights of the calculus samples from batch 2."
#| fig-height: 6
#| fig-width: 7

preservation_plot <- uhplc_calculus_long %>%
  filter(
    !is.na(preservation),
    quant > 0,
    batch == "batch2",
    #!compound %in% c("cbd", "cbn", "cocaine", "thc", "thca-a", "thcva"), # remove compounds not detected
    id != "MB329" # unusually high nicotine and cotinine quantity in batch 1
    ) %>% 
  ggplot(aes(x = preservation, y = quant)) +
    geom_violin(aes(fill = preservation), alpha = 0.6) +
    #geom_jitter(width = 0.2) +
    geom_boxplot(width = 0.2) +
    facet_wrap(~ compound, scales = "free_y", labeller = labeller(compound = compound_names), dir = "v") +
    theme_bw() +
    theme(
      legend.position = "none",
      panel.grid.major.x = element_blank(),
      axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1)
      ) +
    labs(x = "Preservation", y = "Quantity (ng)") +
    scale_fill_viridis_d()

weight_quant_plot <- uhplc_calculus_long %>%
  filter(
    #!is.na(preservation),
    batch == "batch2",
    quant > 0,
    #!compound %in% c("cbd", "cbn", "cocaine", "thc", "thca-a", "thcva"), # remove compounds not detected
    id != "MB329" # unusually high nicotine and cotinine quantity in batch 1
    ) %>% 
  ggplot(aes(x = weight, y = quant)) +
    geom_point(col = viridisLite::viridis(1)) +
    facet_wrap(~ compound, scales = "free_y", labeller = labeller(compound = compound_names), dir = "v") +
    theme_bw() +
    theme(axis.title.y = element_blank()) +
    labs(y = "Quantity (ng)", x = "Weight (mg)")

# weight_quant_plot <- uhplc_data_comb %>%
#   select(sample, contains("batch2")) %>% 
#   select(sample, batch2_weight, contains("calc")) %>% 
#   pivot_longer(
#     -c(sample, batch2_weight),
#     names_to = c("compound", "batch"),
#     names_pattern = "(.*)_(.*)",
#     values_to = c("conc")
#   ) %>% 
#   mutate(compound = str_remove(compound, "_calc")) %>%
#   #remove_missing() %>% 
#   filter(
#     conc > 0,
#     id != "MB329" # unusually high nicotine and cotinine quantity in batch 1
#     ) %>% 
#   ggplot(aes(x = batch2_weight, y = conc)) +
#     geom_point(col = viridisLite::viridis(1)) +
#     facet_wrap(~ compound, scales = "free_y", labeller = labeller(compound = compound_names), dir = "v") +
#     theme_bw() +
#     theme(axis.title.y = element_blank()) +
#     labs(y = "Quantity (ng)", x = "Weight (mg)")

@@

For further statistical analyses, only the UHPLC-MS/MS results from batch 2
were used, as batch 1 had multiple compounds that were not detected in batch 2
and may have been contaminated.

```{r}
#| label: tbl-pearson
#| tbl-cap: "Pearson correlation (*r*) on dichotomous skeletal lesions and compound concentrations (ng/mg) from the second batch. Correlations between pairs of dichotomous variables are removed due to incompatibility with a Pearson correlation. OA = osteoarthritis; VOP = vertebral osteophytosis; SN = Schmorl's nodes; DDD = degenerative disc disease; CO = cribra orbitalia; CMS = chronic maxillary sinusitis; SA = salicylic acid; PN = pipe notches."

@@

```{r}
prettified_names_lower <- sapply(prettified_names, str_to_lower)

corr_tib_out <- corr_tib_long %>%
  mutate(across(where(is.character), plyr::revalue, prettified_names_lower, warn_missing = F))

polycorr_tib_out <- polycorr_tib %>%
  mutate(across(where(is.character), plyr::revalue, prettified_names_lower, warn_missing = F))
```

The strongest point-biserial (Pearson) correlation correlations were a near-perfect
positive correlation between `r cor_statement(corr_tib_out, strength = "strong")`,
and moderate correlations between
`r cor_statement(corr_tib_out, strength = "moderate")` (@tbl-pearson).

Polychoric correlation was conducted on the dichotomised compounds and pathological
conditions, as well as the discretised dental diseases.
Salicylic acid was removed due to its ubiquitous presence in the sample, and is
likely to cause spurious correlations.
Strong correlations were found between `r cor_statement(polycorr_tib_out, strength = "strong")`.
Moderate correlations were found between `r cor_statement(polycorr_tib_out, strength = "moderate")`.
Remaining correlations were weak or absent (@fig-polycorr).
Correlations with age will be depressed because age was largely controlled for
in the sample selection.
